---
title: "Flextables examples"
author: "Alex Thomson - Stats4SD"
date: "04/06/2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = F,
                      warning = FALSE)
```

## Flextable Examples

```{r, echo = FALSE}
library(officedown)
library(officer)
library(flextable)
library(tidyverse)
#setwd("C:/Users/AlexThomson/OneDrive - Statistic For Sustainable Development/June Seminar - tables from R/flextable/examples/")
load("imdb.Rdata")
random_movies <- readRDS("random_movies.rds")
```


## Basic Tables

To create a basic table just need to set up a data frame/tibble with all the data, rows and columns you wish to display. Can simply just assign a new object using `flextable()` if you wish to use all columns in the data frame. Alternatively you can add the argument `col_keys` to select only certain columns from the data.

```{r}
imdb_animation_avgs <- imdb %>%
  filter(type != "videoGame")%>%
  group_by(type, animation)%>%
  summarise(Mean.Votes = mean(numVotes, na.rm = TRUE),
            Mean.Rating = mean(averageRating, na.rm = TRUE),
            Number.Of.Entries = n())

t <- flextable(imdb_animation_avgs) #left assign a new flextable object from your data frame
t
```


### Include a title and a subtitle

Adding a title and subtitle requires a little bit of logic to remember. unfortunately you can't just add them directly like other packages. They are both added by using `add_header_lines` function for which you just need to specify text. However, you want to add the subtitle first as the header line always goes to the very top so you need to start at the bottom of the header and work upwards. 

In this example we have also changed the font size of the main title to make it stand out from the subtitle. In flextable notation `i` means row numbers or names and `j` means column names or numbers. You also want to specify the part of the table as the rows and columns within the header, the body of the table and the footer are treated separately. So in this case the title is row 1 of the header. 

```{r}
top_20_movies <- imdb %>%
  filter(type == "movie" & numVotes > 99999)%>%
  arrange(desc(averageRating))%>%
  select(title, year, averageRating, numVotes)%>%
  slice(1:20)


t <- flextable(top_20_movies)
t <- add_header_lines(t, values = "IMDB Entries with at least 100,000 votes") # add subtitle
t <- add_header_lines(t, values = "Top 20 Movies of all time") # add title
t <- fontsize(t, i = 1, size = 14, part = "header") #increase text size of the title 
t <- autofit(t) # autofit the width of the table and columns
t
```

### change column labels and include number formatting

Changing the labels of columns is quite simple and takes the same format as the `rename` function from dplyr. You simply write `variable.name  = "new column name"`. 

You can set the formatting of a numeric column using `colformat_num` by specifying the column you wish to format (`j = 4` in this example). You can then set the number of decimal places, the big mark, a string for missing values, and also add any prefixing or suffixing if you wish to specify the units directly.

```{r}
t <- flextable(top_20_movies)
t <- set_header_labels(t, 
                       title = "Movie",
                       year = "Year Released",
                       averageRating = "Rating",
                       numVotes = "Total Votes") # written in same fashion as dplyr::rename (variable.name = "new column name")
t <- colformat_num(t, 
                   j = 4, # column number 4
                   digits = 0, # no decimal places
                   big.mark = ",") # use comas when dealing with large numbers
t <- add_header_lines(t, values = "IMDB Entries with at least 100,000 votes") # add subtitle
t <- add_header_lines(t, values = "Top 20 Movies of all time") # add title
t <- fontsize(t, i = 1, size = 14, part = "header") #increase text size of the title
t <- autofit(t)
t
```


### Add borders

Adding borders requires the addition of another package called officer. Using this package you create what is known as a border properties object by setting its colour, style and width e.g. `border_h = fp_border(color="blue", style = "dotted", width = 3)` will create a a thick blue dotted border that can be used within your flextable

```{r}
border_h = fp_border(color="blue", style = "dotted", width = 3)
border_o = fp_border(color = "red", style = "solid", width = 2)

t <- border_inner(t, part = "body", border = border_h)   #SET BORDER FOR EVERY ROW AND COLUMN INSIDE THE TABLE
t <- border_outer(t, border = border_o, part = "body") # SET OUTER BORDER
#t <- border_inner_h(t, part = "body", border = border_h) ONLY HORIZONTAL BORDERS WITHIN THE BODY OF THE TABLE
#t <- border_inner_v(t, part = "body", border = border_h) ONLY SET VERTICAL BORDERS WITHIN THE BODY OF THE TABLE
#t <- hline(t, i = ..., j = ..., part = ..., border = ...)  SET HORIZONTAL BORDERS ON SPECIFIC ROWS/COLUMNS (BOTTOM OF CELL)
#t <- vline(t, i = ..., j = ..., part = ..., border = ...)  SET VERTICAL BORDERS ON SPECIFIC ROWS/COLUMNS (RIGHT OF CELL)
t
```

### Special Features

There are many special features of FlexTable that may not be necessarily needed most of the time but are worth highlighting though they are certainly for someone wishing to do something a bit more advanced.

Most of this revolves around the `compose()` function when used in combination with `as_paragraph()`. This allows you to manually edit text within the table as well as change who the data is visualised including by adding mini bars, line ranges and even adding image files.

```{r}
border_h = fp_border(color="blue")

t <- flextable(random_movies,
               col_keys = c("title", "year", "length", "numVotes", "averageRating", "director")) # use col keys to sepcify columns to plot

t <- add_header_lines(t, values = "IMDB Entries with at least 100,000 votes") # add subtitle

t <- add_header_lines(t, values = "20 Movies") # add title

t <- flextable::align(t, i = 1:2, align = "center", part = "header") # centre align the title and sub title

t <- fontsize(t, size = 12, part = "header") # increase text size of header

t <- colformat_int(t, j = 4, big.mark = ",") # format numbers

t <- flextable::compose(t,
             j = 5,
             value = as_paragraph(
              linerange(value = averageRating)), # default minimum = minum of value and maximum = maximum of value
             part = "body") # turn average rating into a line range visual data representation

t <- flextable::compose(t,
                        j = 3,
                        value = as_paragraph(minibar(value = length, max = max(length))), # default minimum = minum of value and maximum = maximum of value
                        part = "body") # turn length into a mini bar data respresentation


# SET DIRECTORS TO BE IMAGE FILES RATHER THAN THE TEXT
t <- flextable::compose(t,
             i = ~ title == "Chef", # can conditionally set columns and row indexes as well e.g. row where title == "Chef"
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/jon favreau.jpg", width = 2, height = 2))) # additionally specify as_image
# provide a file link to the image
# set width and height of the image
t <- flextable::compose(t,
             i = c(1,8), 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/joel schumacher.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 2, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Jay_Roach.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 3, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Gabor_Csupo.png", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 5, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Ryan_Coogler.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 6, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Alex_Proyas.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 7, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Todd_Phillips.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 9, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Terry_Gilliam.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = c(10, 15), 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Guillermo-del-Toro-2017.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 11, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Eric_Brevig.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 12, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/JUDDPORTRAIT.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 13, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/220px-Dan_Scanlon.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 14, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Nicolas.png", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 16, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Wes_Craven.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 17, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Nancy_Meyers.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 18, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Phillip_Noyce.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 19, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Edward_Zwick.jpg", width = 2, height = 2)))
t <- flextable::compose(t,
             i = 20, 
             j = 6,
            value = as_paragraph(flextable::as_image(src = "directors/Iain_Softley.jpg", width = 2, height = 2)))

t <- width(t, width = 2) # manually set column width

t<-border_inner_h(t, part="body", border = border_h ) # set an interanal border

t
```


### Cotton Farming example

The cotton farming example shown in the seminar showed off how to group rows together with a title for this sub group. This can be done so long as you have a column in your data frame which defines these groups. You can then sue the `as_grouped_data()` function before you create a flextable to set up this grouped data structure which will then create these groupings within the data frame. Think of it like the `group_by` function in dplyr but much more visual. If you look at your data frame after running this function you will see what it does exactly.

```{r, echo = FALSE}
#DATA SET UP CAN IGNORE
CC3_farm <- CC3_farm <- readRDS("CC3_farm.rds")

  X<-CC3_farm%>%
    dplyr::group_by(PartnerCode)%>%
    dplyr::summarise("Mean (Farm)" = round(mean(acre, na.rm = TRUE),2),
                     "Median (Farm)" = round(median(acre, na.rm = TRUE),2),
                     "Mean (Cottton)" = round(mean(cottonacre, na.rm = TRUE),2),
                     "Median (Cotton)" = round(median(cottonacre, na.rm = TRUE),2),
                     "Mean (%Cotton)" = round(mean(cottonprop, na.rm = TRUE),3)*100,
                     "Median (%Cotton)" = round(median(cottonprop, na.rm = TRUE),3)*100,
                     "N" = n())%>%
    dplyr::rename("Sub-group" = PartnerCode)
  
  
  Y<-CC3_farm%>%
    dplyr::group_by(State)%>%
    dplyr::summarise("Mean (Farm)" = round(mean(acre, na.rm = TRUE),2),
                     "Median (Farm)" = round(median(acre, na.rm = TRUE),2),
                     "Mean (Cottton)" = round(mean(cottonacre, na.rm = TRUE),2),
                     "Median (Cotton)" = round(median(cottonacre, na.rm = TRUE),2),
                     "Mean (%Cotton)" = round(mean(cottonprop, na.rm = TRUE),3)*100,
                     "Median (%Cotton)" = round(median(cottonprop, na.rm = TRUE),3)*100,
                     "N" = n())%>%
    dplyr::rename("Sub-group" = State)
  
  
  
  Z<-CC3_farm%>%
    dplyr::mutate(Total = "Total")%>%
    dplyr::group_by(Total)%>%
    dplyr::summarise("Mean (Farm)" = round(mean(acre, na.rm = TRUE),2),
                     "Median (Farm)" = round(median(acre, na.rm = TRUE),2),
                     "Mean (Cottton)" = round(mean(cottonacre, na.rm = TRUE),2),
                     "Median (Cotton)" = round(median(cottonacre, na.rm = TRUE),2),
                     "Mean (%Cotton)" = round(mean(cottonprop, na.rm = TRUE),3)*100,
                     "Median (%Cotton)" = round(median(cottonprop, na.rm = TRUE),3)*100,
                     "N" = n())%>%
    dplyr::rename("Sub-group" = Total)
  A<-rbind(Y,X,Z)
  A$Group <- c(rep("State", 4), rep("Partner", 7), "Overall")
  A <- select(A, Group, `Sub-group`:N)
```

```{r}  
  A <- as_grouped_data(x = A,
                       groups = c("Group")) # group our data on the Group column

  A[is.na(A)] = " " #repalce NAs with an emmpty string
  
  border_h = fp_border(color="gray") # create a border object
  
  t<-flextable(A)
  t<-set_header_labels(t, values = list(Group = "Group", 
                                       `Sub-Group` = "Sub-group",
                                       `Mean (Farm` = str_wrap("Farm Area (Mean)",7), 
                                       `Median (Farm)` = str_wrap("Farm Area (Median)",7),
                                      `Mean (Cottton)` = str_wrap("Cotton Area (Mean)",7),
                                    `Median (Cotton)` = str_wrap("Cotton Area (Median)",7),
                                    `Mean (%Cotton)` = str_wrap("Cotton Area as % of Farm Area (Mean)",12), 
                                    `Median (%Cotton)` = str_wrap("Cotton Area as % of Farm Area (Median)",12),
                                    N = "N"))
  t<-fontsize(t, size = 9, part = "all")
  t<-width(t, width = 0.90)
  t<-add_header_lines(t, values = "Farm Area (Acre), Cotton Area (Acre), Cotton Area as % of Farm Area")
  t <- border_inner_h(t, part="body", border = border_h )
  t
```  

### Modelling Examples

It is additionally not too difficult to create modelling tables using flextable. You could combine it with the very useful function `tidy()` from the broom packages which turns model summaries into a convenient tidy data fame that can be used as a table. This tidy function can be used for numerous different types of statistical models.

Additionally however, with any glm, lm and some statistical tests you can directly pipe in the `as_flextable()` function and this will automatically create a model table that includes significance codes and footnotes on various model characteristics such as the R-squared and F-statistic.
 
```{r}
movies <- imdb%>%
  filter(type == "movie")%>%
  mutate(num_Votes_10000 = numVotes/10000)

model1 <- aov(averageRating ~ length + num_Votes_10000 + animation, movies)

m1 <- broom::tidy(model1)

t <- flextable(m1)
t <- set_header_labels(t, values = list(
  term = "Term",
  df = "Df",
  sumsq = "SS",
  meansq = "MSS",
  statistic = "F - Statistic",
  p.value = "P.Value"
))
t <- colformat_double(t, j = c(3:6), digits = 2, na_str = " ")
t <- autofit(t)
t
```

```{r}
lm(averageRating ~ length + num_Votes_10000 + animation, movies) %>%
  as_flextable()
```

```{r}
t.test(averageRating ~ fantasy, data = movies)%>%
  as_flextable()
```