---
title: "KableExtra"
author: "Alex Thomson - Stats4SD"
date: "05/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning  = FALSE)
```

```{r, include = FALSE}
library(kableExtra)
library(tidyverse)
setwd("C:/Users/AlexThomson/OneDrive - Statistic For Sustainable Development/June Seminar - tables from R/kable/")
load("imdb.Rdata")
CC3_farm <- readRDS("CC3_farm.rds")
```

## Basic Tables

One can create an extremely basic table using just the `kable()` function and no formatting. This table feature no formatting and not incorporate any borders between cells. 

You do not have to specify the `format = ` argument, if you do not the table will be created regardless of your output type (word, html, etc.). Specifying this argument is mostly required for additional changes to the appearance of the table when using the kableExtra package. In the case below, we have specified a html output, if we attempted to knit this document to word or pdf the table would be created.

```{r}
imdb_animation_avgs <- imdb %>%
  filter(type != "videoGame")%>%
  group_by(type, animation)%>%
  summarise(Mean.Votes = mean(numVotes, na.rm = TRUE),
            Mean.Rating = mean(averageRating, na.rm = TRUE),
            Number.Of.Entries = n())

knitr::kable(imdb_animation_avgs, "html") #basic table with no editing of appearance
```

### include a title and change column names

Unlike other packages, you can set many formatting options within the call to `kable()` this includes arguments for the number of digits for numeric columns, row names, column names, column alignment and a caption. As well as many arguments which can be passed to the `format.args = ` option.

Using the kableExtra package, we are then able to create a much nicer and cleaner looking output for our table. This package is mainly designed to improve the appearance for tables in html and pdf, with different options available for these. You can pipe in the `kable_styling()` function with no arguments which will provide a default sleeker looking appearance to the table. The appearance depends on the format specified in the kable function. The example below focuses on html.

```{r}
top_20_movies <- imdb %>%
  filter(type == "movie" & numVotes > 99999)%>%
  arrange(desc(averageRating))%>%
  select(title, year, averageRating, numVotes)%>%
  slice(1:20)

knitr::kable(top_20_movies,
             "html",
             col.names = c("Movie", "Year", "Rating", "Total Votes"), #provide a string of column names
             caption = "IMDB Entries with at least 100,000 votes")%>% # provide a caption (title)
  kable_styling()
```

### change number formatting and collapse rows

Additionally within the `kable()` function you can change number formatting by using the digits argument. You can do this by providing vector equal to the number of columns with the number of digits for each column if you wish to have columns with different arguments. Alternatively you can just provide a single number for a global option applicable to the whole table.

Similarly you can provide a string to the align argument to specify the column alignments, with "l" for left, "r" for right and "c" for centre. For instance if you write a string of "clc" then the first column will be centre aligned, the second left aligned and the third centre aligned.

You can use the `collapse_rows()` function to group same values in columns into multiple row cells. The main argument here is to provide the column you wish to apply this to. You can also specify the `valign = ` argument by providing a string saying either top, middle, or bottom. This specifies the vertical alignment of the text in the grouped cell.

```{r}
knitr::kable(imdb_animation_avgs,
             "html",
             col.names = gsub("[.]", " ", names(imdb_animation_avgs)), # specify column names, can use a formatting function instead of a string if you so wish
             caption = "Popularity and Quality of Animation",
             digits = c(0,0,0,2,0), # specify the digits for all columns
             format.args = list(big.mark = ","), # include a big mark for numbers above 1000
             align = "llrrr")%>% #RULER OF OMICRON PERSEI 8
            #provide a character string to set column aligmnets
  collapse_rows(columns = 1, valign = "middle")%>% # group 1st coulumn by the same value and align text in the middle of the cell
  kable_styling()
```

### add a footnote

You can also add footnotes using the `footnote()` function. You can create different types of footnotes with different symbols using different arguments. `general = ` will create an overall footnote with no symbol attached, `number = ` will create a numeric symbol footnote, `alphabet = ` will create an alphabetic footnote and `symbol = ` will create footnotes using special symbols. Special symbols can also be manually specified.

These footnotes will not specify to a particular cell/title in the table unfortunately, something you can do with gt. In order to do this you need to edit the particular cell beforehand, or if it is a column name then you can do so within the `col.names` argument. You specify which footnote this applies to by using the `footnote_marker_***()` function. This specifies which list of footnotes to use and which symbol in the list to use. For example, using `footnote_marker_alphabet(1)` will put an "a" after the value, `footnote_marker_alphabet(2)` will put a "b" after the value and so on and so on.

If you are ever using these footnote marker functions, then within your call to kable, you have to include the `escape = FALSE` argument. This allows the function escape special characters when converting to the table. This will avoid any syntax errors created by these symbols.

```{r}
top_20_movies[5,3] <- paste(top_20_movies[5,3], footnote_marker_symbol(1)) #sepcify the first sybmbol (* default)

knitr::kable(top_20_movies,
             "html",
             col.names = c(paste("Movie", footnote_marker_alphabet(1)), "Year", "Rating", "Total Votes"), #include and "a" agter the column name
             caption = "IMDB Entries with at least 100,000 votes",
             escape = F)%>% # ensure that html code is correctly honoured
  kable_styling()%>%
    footnote(general = "Here is a general comments of the table. ", #use vector of character strings to write multiple footnotes at once 
           number = c("Footnote 1; ", "Footnote 2; "),
           alphabet = c("Footnote A; ", "Footnote B; "),
           symbol = c("Footnote Symbol 1; ", "Footnote Symbol 2"))
```

## Special Features

### text and cell colours

This is quite a complicated example but shows some of the extensions one can do within kable. This using `cell_spec()` to specify the format of a cell before sending the data frame to kable. The function will provide LaTex of html code formatting to the cells of your data frame including text colour, background colours, italic text, bold text etc. You can even use the `spec_color()` function which will apply a continuous colour gradient to your cells.

```{r}
iris[1:10, ] %>%
mutate_if(is.numeric, function(x) {
cell_spec(x, "html", bold = T, color = spec_color(x, end = 0.9), # mutate the cells of the data frame directly, cell_spec will add in html or Latex coding which kable will be able to read
font_size = spec_font_size(x)) # specify font size, changes relative to range of values in the column in this case. A continous scale of font size and colour
}) %>%
mutate(Species = cell_spec(
Species, "html", color = "white", bold = T,
background = spec_color(1:10, end = 0.9, option = "C", direction = -1) #specify a background contninous colour gradient
)) %>%
kable("html", escape = F, booktabs = T, linesep = "", align = "c")%>% # remember the escape argument to make sure html code is read properly
  kable_styling()
```

### scroll box

There are many html only features that can be used within kableExtra that could especially be useful if ever wanting to create tables in a shiny application such as one of our data monitoring platforms.

This is done using the `scroll_box()` function in which you specify a character string defining the width and height of the box using either pixels or a percentage.

```{r}
top_100_movies <- imdb %>%
  filter(type == "movie" & numVotes > 99999)%>%
  arrange(desc(averageRating))%>%
  slice(1:100)

kable(top_100_movies,
      "html")%>%
  kable_styling()%>%
  scroll_box(width = "500px", height = "400px") # specify size of the scroll box
```

## Cotton Connect Example

The cotton connect example shown within the seminar is fairly straightforward in comparison to the other packages despite kable being a more technical option. This is partly due to us being able to manually create our row groupings using `pack_rows()` rather than requiring an existing grouping column. Therefore in this case we did not need to create an additional column beforehand or use an additional function to directly group the data before we even made a table. 

This function requires you to provide a kable for the grouping and then to specify the first and last rows within this new grouping.

```{r, echo=FALSE}
  X<-CC3_farm%>%
    dplyr::group_by(PartnerCode)%>%
    dplyr::summarise("Mean (Farm)" = round(mean(acre, na.rm = TRUE),2),
                     "Median (Farm)" = round(median(acre, na.rm = TRUE),2),
                     "Mean (Cottton)" = round(mean(cottonacre, na.rm = TRUE),2),
                     "Median (Cotton)" = round(median(cottonacre, na.rm = TRUE),2),
                     "Mean (%Cotton)" = round(mean(cottonprop, na.rm = TRUE),3)*100,
                     "Median (%Cotton)" = round(median(cottonprop, na.rm = TRUE),3)*100,
                     "N" = n())%>%
    dplyr::rename("Sub-group" = PartnerCode)
  
  
  Y<-CC3_farm%>%
    dplyr::group_by(State)%>%
    dplyr::summarise("Mean (Farm)" = round(mean(acre, na.rm = TRUE),2),
                     "Median (Farm)" = round(median(acre, na.rm = TRUE),2),
                     "Mean (Cottton)" = round(mean(cottonacre, na.rm = TRUE),2),
                     "Median (Cotton)" = round(median(cottonacre, na.rm = TRUE),2),
                     "Mean (%Cotton)" = round(mean(cottonprop, na.rm = TRUE),3)*100,
                     "Median (%Cotton)" = round(median(cottonprop, na.rm = TRUE),3)*100,
                     "N" = n())%>%
    dplyr::rename("Sub-group" = State)
  
  
  
  Z<-CC3_farm%>%
    dplyr::mutate(Total = "Total")%>%
    dplyr::group_by(Total)%>%
    dplyr::summarise("Mean (Farm)" = round(mean(acre, na.rm = TRUE),2),
                     "Median (Farm)" = round(median(acre, na.rm = TRUE),2),
                     "Mean (Cottton)" = round(mean(cottonacre, na.rm = TRUE),2),
                     "Median (Cotton)" = round(median(cottonacre, na.rm = TRUE),2),
                     "Mean (%Cotton)" = round(mean(cottonprop, na.rm = TRUE),3)*100,
                     "Median (%Cotton)" = round(median(cottonprop, na.rm = TRUE),3)*100,
                     "N" = n())%>%
    dplyr::rename("Sub-group" = Total)
  A<-rbind(Y,X,Z)
```
  
```{r}  
knitr::kable(A,
             "html",
             col.names = c("Sub-group", 
                           "Farm Area (Mean)",
                           "Farm Area (Median)", 
                           "Cotton Area (Mean)",
                           "Cotton Area (Median)",
                           "Cotton Area as % of Farm Area (Mean)",
                           "Cotton Area (Median)",
                           "N"),
             caption = "Farm Area (Acre), Cotton Area (Acre), Cotton Area as % of Farm Area")%>%
  kable_styling(font_size = 9)%>%
  pack_rows(group_label = "State", 1, 4)%>% # include a label, the first and last rows of the grouping
  pack_rows(group_label = "Partner",5,11)%>%
  pack_rows(group_label = "Overall",12,12)
```

## Modelling Examples

For creating model tables, again the solution is to create a tidy model data frame using the broom package and then using kable on this data frame. You can edit this table in any way you wish. One extra thing to note is that you can use the code `opts <- options(knitr.kable.NA = "")` to specify an empty string for missing data instead of "NA".

```{r}
movies <- imdb%>%
  filter(type == "movie")%>%
  mutate(num_Votes_10000 = numVotes/10000)

model1 <- aov(averageRating ~ length + num_Votes_10000 + animation, movies)

m1 <- broom::tidy(model1) # create a tidy model table

opts <- options(knitr.kable.NA = "") # set NA values to an empty string

kable(m1,
      "html",
      digits = c(0,0,0,0,1,5))%>%
  kable_styling()
```

```{r}
lm(averageRating ~ length + num_Votes_10000 + animation, movies) %>%
  broom::tidy()%>%
  kable("html", 
        digits = c(0,3,3,5))%>%
  kable_styling()
```